Generating Performance Improvement Suggestions by using Cross-Organizational Process Mining

### 2015-11-21 4_methodology/methodology.tex
Remove itemize in discussion
Shortening in progress

\subsubsection{Performance Indicator Clustering}
\label{subsubsec:performance-indicator-clustering}
Clustering is based on the idea of collecting the set of observations into clusters so that observations within the same cluster are similar whereas the observations from different clusters are dissimilar. In this study, clustering is used to gather organizations based on their performance indicator data. In this research, random initialization based \textit{k-means++} approach from the study of Arthur and Vassilvitskii \cite{arthur2007} is used to cluster organizations. Since the number of clusters are not known priori, k-means clustering is applied starting \textit{k} from 1 to the number of organizations. For each clustering with different number of clusters, \textit{Sum of Squared Error (SSE)} values are plotted and user is asked to select the appropriate cluster size. For the selected cluster size, clustering related information is used to generate recommendations in the further steps. Resulting cluster analysis data is formulated as follows:

\theoremstyle{definition}
\begin{definition}
Cluster Analysis Data is a tuple $(k, Assignments, Cluster\ Centroids)$ where
\begin{enumerate}
  \item $k$ is the number of clusters,
  \item $Assignments$ is a set of tuple $(Organization_{i}, Cluster_{j})$ where $i$ is identifier for organization and $j \leq k$ is identifier for cluster,
  \item $Cluster\ Centroids$ is a set of tuple $(Cluster_{j}, Type, A_{start}, A_{end}, Value)$ where
    \begin{enumerate}
      \item $Type$ is performance indicator type which is $Average$ or $StandardDev$,
      \item $A_{start}$ and $A_{end}$ are starting and ending points of performance indicator,
      \item $Value$ is the actual value of performance indicator,
      \item $Cluster\ Centroids_{j}$ is a function that returns a set of $Centroid$ which is a tuple $(Type, A_{start}, A_{end}, Value)$ for $Cluster_{j}$.
    \end{enumerate} 
\end{enumerate}
\end{definition}

@@

\subsection{Mismatch Pattern Analysis}
\label{subsec:mismatch-pattern-analysis}
In order to learn from other organizations, it is necessary to spot the differences between process models of different organizations. In this phase, differences between process models will be revealed by the mismatch patterns which are defined by Dijkman \cite{dijkman2007mismatch}. In the process mining stage, process models are mined to create Workflow nets; however, mismatch pattern definitions are closer to business environment and it is more suitable spot patterns in BPMN notation. Thus, mined process models are converted to BPMN models and then mismatch pattern analysis is applied in order to locate the differences between two process models or two clusters of process models. In addition, since performance indicators are calculated based on a starting and ending point in the process model, the same approach is applied to locate mismatch patterns. In other words, differences of process models are located through a starting activity to an ending activity. With this aim, each mismatch pattern and its analyzers are defined by extending the following definitions:
\theoremstyle{definition}
    \begin{definition}
    Mismatch Pattern is a tuple ${Mismatch\ Patten} = (O_{1}, O_{2}, ExtensionData, A_{start}, A_{end}) $ where 
    \begin{enumerate}
      \item $O_{1}$ is the first organization and \item $O_{2}$ is the second organization in between the pattern occurs,
      \item $ExtensionData$ is a set of tuples where mismatch related information is recorded, 
      \item $A_{start}$ and $A_{end}$ are starting and ending points to check mismatch patterns.
    \end{enumerate}
    \end{definition}

    \theoremstyle{definition}
    \begin{definition}
    Mismatch Pattern Analyzer is a function $MismatchPatternAnalyzer(O_{1}, O_{2}, A_{start}, A_{end})$ and it returns a set of \textit{Mismatch Pattern} for the organization $O_{1}$ compared to $O_{2}$ for the activities between $A_{start}$ and $A_{end}$.
    \end{definition}

@@

\subsection{Genarating Suggestions/Recommendations for Performance Improvement}
\label{subsec:recommendation-generation}
Recommendation generation stage in the methodology is the final and core stage where all information retrieved from the event logs until now are utilized. In this study, idea of recommendation is based on providing a set of mismatch patterns for each organization so that they can enhance their processes. These mismatch patterns are generated by comparing the process models of other organizations, particularly those that are performing better in terms of their performance indicator values. In addition, clustering of organizations is undertaken to generalize the way of identifying which organizations perform better in this environment. Recommendation idea and recommendation generation function is defined as following:
\theoremstyle{definition}
\begin{definition}
Recommendation is a tuple ${Recommendation} = (O, A_{start}, A_{end}, Mismatch\ Patterns) $ where 
  \begin{enumerate}
    \item $O$ is identifier for organization,
    \item $A_{start}$ and $A_{end}$ are starting and ending activities in between the recommendations are checked,
    \item $Mismatch\ Patterns$ is collection of mismatch patterns.
  \end{enumerate}
\end{definition}

@@

This stage of methodology can be visualized as gathering inputs of mismatch patterns data and cluster analysis data in Figure~\ref{fig:recommendation-generation-blackox}. In addition, a performance difference threshold is gathered from the user to specify how much difference between clusters is necessary to check for mismatch patterns and generate recommendations. This stage generates recommendation data for each organization that contains a set of mismatch patterns.
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{4_methodology/recommendation-generation-blackbox}
  \caption{Recommendation Generation Stage as Black-box}
  \label{fig:recommendation-generation-blackox}
\end{figure}

>>>>

\subsubsection{Performance Indicator Clustering}
\label{subsubsec:performance-indicator-clustering}
Clustering is based on the idea of collecting the set of observations into clusters so that observations within the same cluster are similar whereas the observations from different clusters are dissimilar. In this study, clustering is used to gather organizations based on their performance indicator data. In this research, random initialization based \textit{k-means++} approach from the study of Arthur and Vassilvitskii \cite{arthur2007} is used to cluster organizations. Since the number of clusters are not known priori, k-means clustering is applied starting \textit{k} from 1 to the number of organizations. For each clustering with different number of clusters, \textit{Sum of Squared Error (SSE)} values are plotted and user is asked to select the appropriate cluster size. For the selected cluster size, clustering related information is used to generate recommendations in the further steps. Resulting cluster analysis data is formulated as follows:
\theoremstyle{definition}
\begin{definition}
Cluster Analysis Data is a tuple $(k, Assignments, Cluster\ Centroids)$ where
\begin{enumerate}
  \item $k$ is the number of clusters,
  \item $Assignments$ is a set of tuple $(Organization_{i}, Cluster_{j})$ where $i$ is identifier for organization and $j \leq k$ is identifier for cluster,
  \item $Cluster\ Centroids$ is a set of tuple $(Cluster_{j}, Type, A_{start}, A_{end}, Value)$ where
    \begin{enumerate}
      \item $Type$ is performance indicator type which is $Average$ or $StandardDev$,
      \item $A_{start}$ and $A_{end}$ are starting and ending points of performance indicator,
      \item $Value$ is the actual value of performance indicator,
      \item $Cluster\ Centroids_{j}$ is a function that returns a set of $Centroid$ which is a tuple $(Type, A_{start}, A_{end}, Value)$ for $Cluster_{j}$.
    \end{enumerate} 
\end{enumerate}
\end{definition}

@@

\subsection{Mismatch Pattern Analysis}
\label{subsec:mismatch-pattern-analysis}
In order to learn from other organizations, it is necessary to spot the differences between process models of different organizations. In this phase, differences between process models will be revealed by the mismatch patterns which are defined by Dijkman \cite{dijkman2007mismatch}. Since performance indicators are calculated based on a starting and ending point in the process model, the same approach is applied to locate mismatch patterns. In other words, differences of process models are located through a starting activity to an ending activity. With this aim, each mismatch pattern and its analyzers are defined by extending the following definitions:
\theoremstyle{definition}
    \begin{definition}
    Mismatch Pattern is a tuple ${Mismatch\ Patten} = (O_{1}, O_{2}, ExtensionData, A_{start}, A_{end}) $ where 
    \begin{enumerate}
      \item $O_{1}$ is the first organization and \item $O_{2}$ is the second organization in between the pattern occurs,
      \item $ExtensionData$ is a set of tuples where mismatch related information is recorded, 
      \item $A_{start}$ and $A_{end}$ are starting and ending points to check mismatch patterns.
    \end{enumerate}
    \end{definition}
\theoremstyle{definition}

\begin{definition}
    Mismatch Pattern Analyzer is a function $MismatchPatternAnalyzer(O_{1}, O_{2}, A_{start}, A_{end})$ and it returns a set of \textit{Mismatch Pattern} for the organization $O_{1}$ compared to $O_{2}$ for the activities between $A_{start}$ and $A_{end}$.
\end{definition}

@@

\subsection{Genarating Suggestions/Recommendations for Performance Improvement}
\label{subsec:recommendation-generation}
\theoremstyle{definition}
Recommendation generation stage in the methodology is the final and core stage where all information retrieved from the event logs until now are utilized. In this study, idea of recommendation is based on providing a set of mismatch patterns for each organization so that they can enhance their processes. These mismatch patterns are generated by comparing the process models of other organizations, particularly those that are performing better in terms of their performance indicator values. Recommendation idea and recommendation generation function is defined as following:
\begin{definition}
Recommendation is a tuple ${Recommendation} = (O, A_{start}, A_{end}, Mismatch\ Patterns) $ where 
  \begin{enumerate}
    \item $O$ is identifier for organization,
    \item $A_{start}$ and $A_{end}$ are starting and ending activities in between the recommendations are checked,
    \item $Mismatch\ Patterns$ is collection of mismatch patterns.
  \end{enumerate}
\end{definition}

@@

This stage of methodology can be visualized as gathering inputs of mismatch patterns data and cluster analysis data in Figure~\ref{fig:recommendation-generation-blackox}.
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{4_methodology/recommendation-generation-blackbox}
  \caption{Recommendation Generation Stage as Black-box}
  \label{fig:recommendation-generation-blackox}
\end{figure}
